---
title: "DATA_607_Week9_Homework_Bikram"
author: "Bikram Barua"
date: "10/22/2021"
output: rmdformats::readthedown
---

### Assignment – Web APIs

### The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis.
### Task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.
<br>

##### Load the libraries
```{r warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)
library(listviewer)
library(kableExtra)
library(tidyverse)
library(purrr)
library(htmltools)
```

<br>

# New York Times- Top(Technology) Stories API

##### Call the API
##### View the JSON data response attributes
```{r warning=FALSE, message=FALSE}

tech_response <- GET("https://api.nytimes.com/svc/topstories/v2/technology.json?api-key=97KSRsEE7aPgeEDTu3tZT4sVYND9NofH")

raw_tech_json = fromJSON(rawToChar(tech_response$content))

names(raw_tech_json)

```

<br>

# Interactive `View` of the response
```{r}


# Interactive JSON results viewer
listviewer::jsonedit(raw_tech_json$results, height = "400px", mode = "view")

```

<br>

# Format date columns
```{r}


publish_date <- as.POSIXct( substr(raw_tech_json$results$published_date,1,10)  , format = "%Y-%m-%d")
raw_tech_json$results$fmt_published_date <- format(publish_date, "%m/%d/%Y")


create_date <- as.POSIXct( substr(raw_tech_json$results$created_date,1,10)  , format = "%Y-%m-%d")
raw_tech_json$results$fmt_created_date <- format(create_date, "%m/%d/%Y")


```


# Converting JSON data into R dataframe
```{r warning=FALSE, message=FALSE}

# Get the Facet information by each story (row_n)
get_facet_data <- function(row_n) {
  purrr::pluck(raw_tech_json, "results", "des_facet", row_n)
}

# create the dataframe and tidy it up
tech_stories_df <- pluck(raw_tech_json, "results") %>%
  # convert to tibble
  as_tibble() %>%
  # select columns of interest
  select(title, abstract, fmt_published_date, fmt_created_date ) %>%
  mutate(data = map(row_number(), get_facet_data)) %>% 
  unnest_wider(data) %>%
  mutate(published_date =    )

names(tech_stories_df) <- c('Title', 'Story Abstract', 'Published Date', 'Created Date', 'Facet 1', 'Facet 2', 'Facet 3', 'Facet 4', 
                       'Facet 5', 'Facet 6', 'Facet 7', 'Facet 8', 'Facet 9', 'Facet 10', 'Facet 11', 'Facet 12', 'Facet 13' )


knitr::kable(tech_stories_df, format = "html")
```

