---
title: "DATA_607_Project1"
author: "Bikram Barua"
date: "9/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 607 Project1


##### Let's load the required libraries in R for data analysis
```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)
library(downloader)
library(stringr)
library(htmlTable)
library(tidyverse)
library(janitor)

library(sqldf)
library(RODBC)
library(DBI)
library(odbc)
```

<br>
##### Download and read the txt file to a dataframe object
##### Filter the empty rows

```{r message=FALSE, echo=FALSE, warning=FALSE}

getwd() # Path where the file will be downloaded
url <- "https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/Project1_Data1.txt"
download.file(url, "Project_Data.txt")


data <- read.delim(file = "Project_Data.txt", header = FALSE, sep = "|", dec = ".")
# Filter empty rows
data1 <- subset( data,V2 != "") 

# Extract Odd rows from dataframe
odd_rows <- seq_len(nrow(data1)) %% 2
# Drop even rows
data_odd_rows <- data1[odd_rows == 1, ]

# Now Drop odd rows, create a dataframe of even rows
data_even_rows <- data1[odd_rows == 0, ]
subset_even_rows <- subset(data_even_rows, select= c(V1, V2))

# Rename column names
colnames(subset_even_rows) <- c("State","Rating")
colnames(data_odd_rows) <- c("Player_Num","Player_Name","Total_Pts","Round_1","Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7")

# Combine the two data frames
data_rows <- cbind(data_odd_rows,subset_even_rows)

# String manipulation on column to extract Player Rating
rating_data <- str_extract_all(data_rows$Rating,boundary("word"), simplify = T)

# Create a new column for Player's Pre-Rating and add it to the existing dataframe
new_data_rows <- data_rows
new_data_rows <- cbind(data_rows, Player_PreRating = rating_data[,3])


```


<br>

##### Parse the Game Round information to isolate the opponent Player_Nums for further analysis

```{r}

round1_data <- str_extract_all(new_data_rows$Round_1,boundary("word"), simplify = T)
round2_data <- str_extract_all(new_data_rows$Round_2,boundary("word"), simplify = T)
round3_data <- str_extract_all(new_data_rows$Round_3,boundary("word"), simplify = T)
round4_data <- str_extract_all(new_data_rows$Round_4,boundary("word"), simplify = T)
round5_data <- str_extract_all(new_data_rows$Round_5,boundary("word"), simplify = T)
round6_data <- str_extract_all(new_data_rows$Round_6,boundary("word"), simplify = T)
round7_data <- str_extract_all(new_data_rows$Round_7,boundary("word"), simplify = T)

new_data_rows <- cbind(new_data_rows, Round1_Opponent = round1_data[,2], Round2_Opponent = round2_data[,2], 
        Round3_Opponent = round3_data[,2], Round4_Opponent = round4_data[,2], Round5_Opponent = round5_data[,2],
        Round6_Opponent = round6_data[,2], Round7_Opponent = round7_data[,2])

subset_db_rows <- subset(new_data_rows, select= c(Player_Num, Player_Name, Total_Pts, Player_PreRating, Round1_Opponent,Round2_Opponent,Round3_Opponent,Round4_Opponent,Round5_Opponent,Round6_Opponent,Round7_Opponent))

## Substitute blank Opponent Player_Num to 99999
subset_db_rows[subset_db_rows == ""] <- 99999

db_rows <- subset_db_rows

## Filter the Rating digits, removing the P?? numbers
PreRating <- str_extract_all(subset_db_rows$Player_PreRating, "^[:digit:]+")

PreRating.df <- as.data.frame(t(PreRating))
T_PreRating <- t(PreRating.df)
 
## Add a new column with just Pre_Rating
db_rows$Pre_Rating <- T_PreRating    

## First coerce the data.frame to all-character
db_rows_coerce = data.frame(lapply(db_rows, as.character), stringsAsFactors=FALSE)


```


<br>

##### Create dataframes to help compute 'Average Pre Chess Rating of Opponents'
##### Convert the character value in the dataframe to Integer to help compute the mean
```{r}
player_opponents.df <- subset(db_rows_coerce, select = c(Player_Num, Pre_Rating, Round1_Opponent, Round2_Opponent, Round3_Opponent, Round4_Opponent, Round5_Opponent, Round6_Opponent, Round7_Opponent) )

data1 <- player_opponents.df

dat <- as.data.frame(sapply(data1, as.numeric)) #<- sapply is here

dat_num <- dat

```

##### Loop thru the dataframe to get the Opponent Rating and calculate the Mean_Opponent_Rating
##### Bypass the byes, forfeit in mean calculation
##### Combine the dataset with the original dataframe with Player Name and other information
```{r}
for(i in 1:nrow(dat_num)) {       # for-loop over rows
   count <- 0
   sum <- 0
   j <- 3
   for(j in 3:ncol(dat_num)) {  
      dat_num[i, j] <- dat_num[dat_num[i, j], 2] # Replace Opponent player_num with their Pre_Rating
      if (is.na(dat_num[i,j]) == FALSE) {
        count <- count + 1
        sum <- sum + dat_num[i,j]
      }
   }
   dat_num[i,10] <-   format(round(sum/count, 2), nsmall = 2) 
   dat_num[i,11] <- count
}

## Combine the original dataframe, map with the ratings
combined_df <- cbind(data_rows, dat_num[ ,2], dat_num[ ,10], dat_num[ ,11])              
colnames(combined_df) <- c("Player_Num","Player_Name", "Total_Pts",  "Round_1","Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7", "Temp","State","All Rate Info", "Pre_Rating","Opponent_Mean_Rating", "Num_Games_Played")

result_df <- subset(combined_df, select= c(Player_Num, Player_Name, State, Total_Pts, Pre_Rating, Opponent_Mean_Rating), row_number=FALSE)
htmlTable(result_df)
```


##### Create Player dataset who played all the scheduled games
##### Create Player dataset who had atleast one unplayed game
```{r}

all_games_played_df <- subset(combined_df, select= c(Player_Num, Player_Name, Total_Pts, Pre_Rating, Opponent_Mean_Rating,Num_Games_Played), row_number=FALSE)

## Players who played 7 Rounds
max_games_played_df <- filter(all_games_played_df, all_games_played_df$Num_Games_Played == 7)
htmlTable(max_games_played_df)

## Players who had one or more unplayed games 
games_unplayed_df <- filter(all_games_played_df, all_games_played_df$Num_Games_Played < 7)
htmlTable(games_unplayed_df)



```

##### Sort the Players by their Pre-Ratings, Points

```{r}
# Sort by Points and Pre_Ratings
sorted_df <- result_df[order(result_df$Total_Pts, result_df$Pre_Rating, decreasing = TRUE),]

htmlTable(sorted_df, rnames = FALSE)

```

#### **GARY HUA scored the most points relative to his expected result**


##### Create the CSV file
```{r}
## Create the .CSV file from the dataframe
getwd() # Path where the file will be downloaded
write.csv(result_df,"Chess_Project_Result.csv", row.names = FALSE)


```


