---
title: "DATA_607_Project2_2"
author: "Bikram Barua"
date: "9/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project 2 (Global Child Mortality Rates data)
This is a dataset from Unicef with mortality rates under 5 years across different countries from 1950 to 2015.
Data Reference: <https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/unicef-u5mr.csv>


##### Let's load the required libraries in R for data analysis
```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)
library(downloader)
library(stringr)
library(htmlTable)
library(tidyverse)
library(reshape2)
```

<br>

##### Download and read the CSV file to a dataframe object  
##### View the data  

```{r message=FALSE, echo=TRUE, warning=FALSE}

# Import the data from the csv file
unicef_data <- read.csv("https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/unicef-u5mr.csv", stringsAsFactors = FALSE)

head(unicef_data, 5) 

```


##### Clean up the column names

```{r}
names(unicef_data) <- gsub("U5MR.", "", names(unicef_data))

```


##### Remove the first column of the data frame

```{r}
subset_unicef_data <- select( unicef_data, -1)

# Assign row averages to new variable named row_mean
subset_unicef_data$row_mean <- round(rowMeans(subset_unicef_data, na.rm=TRUE),2)

# Add the first column back to the dataframe 
bind_unicef_data <- cbind(select(unicef_data, 1), subset_unicef_data)

# Sort the dataframe by the mean of the mortality rate 
highest_child_mortality <-  arrange(bind_unicef_data, desc(row_mean)) %>% head(5) 

# Display the countries with highest rates
htmlTable( select( highest_child_mortality, c('CountryName', 'row_mean')))

```

#### Lets do some analysis on the top 5 countries, to find any relationship.
#### Subset last 5 yrs from the dataset for the top 5 countries from (2011 - 2015)
#### Transform the dataset from wide format to long format
```{r warning=FALSE, message=FALSE}
subset_child_mortality <-  select( highest_child_mortality, c('CountryName', '2011','2012','2013', '2014','2015'))

subset_child_mortality1 <- subset_child_mortality %>%
    gather("year", "cases", "2011":"2012":"2013":"2014":"2015")

# Plot the country line graph by the year to identify trend 
ggplot(subset_child_mortality1, aes(CountryName,cases, group=year)) + 
   geom_line(aes(color=year))+
 geom_point(aes(shape=year))

```
The mortality rates consistently decreased from 2011 - 2015 for all the countries.

<br>

#### For the last 5 years, find mean, standard deviation, variance of the cases

```{r warning=FALSE, message=FALSE}
mean_child_mortality <- subset_child_mortality1 %>% group_by(CountryName) %>%  summarise(avg_deaths = mean(cases, na.rm=TRUE)) 

sd_child_mortality <- subset_child_mortality1 %>% group_by(CountryName) %>%  summarise(sd_deaths = sd(cases, na.rm=TRUE))

min_child_mortality <- subset_child_mortality1 %>% group_by(CountryName) %>%  summarise(min_deaths = min(cases, na.rm=TRUE))

max_child_mortality <- subset_child_mortality1 %>% group_by(CountryName) %>%  summarise(max_deaths = max(cases, na.rm=TRUE))

variance_child_mortality <- subset_child_mortality1 %>% group_by(CountryName) %>%  summarise(variance_deaths = var(cases, na.rm=TRUE))

child_mortality_stats <- mean_child_mortality %>% inner_join(sd_child_mortality, by = 'CountryName') %>% inner_join(min_child_mortality, by = 'CountryName') %>% inner_join(max_child_mortality, by = 'CountryName') %>% inner_join(variance_child_mortality, by = 'CountryName') 

arrange(child_mortality_stats, desc(avg_deaths)) 
  

```


The trend of different statistical values is consistent across countries.

```{r warning=FALSE, message=FALSE}
child_mortality_stats1 <- child_mortality_stats 

library(reshape2)
#melt data frame into long format
df <- melt(child_mortality_stats1 ,  id.vars = 'CountryName')

# Everything on the same plot
ggplot(df, aes(CountryName,value, col=variable)) + 
  geom_line(aes(color=variable))+
  geom_point()  

```


#### Load another dataset with Continent/Country mapping data 

```{r}

# Import the data from the csv file
continent_data <- read.csv("https://pkgstore.datahub.io/JohnSnowLabs/country-and-continent-codes-list/country-and-continent-codes-list-csv_csv/data/b7876b7f496677669644f3d1069d3121/country-and-continent-codes-list-csv_csv.csv", stringsAsFactors = FALSE)

head(continent_data, 5)

continent_data$Country_Name <- gsub(",.*$", "", continent_data$Country_Name)

continent_data = continent_data %>% dplyr::mutate(CountryName = gsub(",.*$", "", continent_data$Country_Name))

```

#### Join the two datasets
#### Group the data by Continent, summarize and sort them by their cumulative totals

```{r}
joined_df <- left_join( continent_data,bind_unicef_data, by = "CountryName")

sum_cont_df <- joined_df  %>% select( Continent_Name, Continent_Code, row_mean) %>% group_by( Continent_Code)  %>% mutate(sum_mean = sum(row_mean, na.rm=TRUE)) %>% summarize(sum_mean = max(sum_mean)) %>% arrange(desc(sum_mean))


htmlTable(sum_cont_df)
```


#### Join the top 5 counties dataset with the continent dataset

```{r}
top_cont_country_df <- inner_join( continent_data,subset_child_mortality, by = "CountryName")
top_cont_country_df <- top_cont_country_df  %>% select( Continent_Name, Continent_Code, CountryName) 
htmlTable(top_cont_country_df)
```

This is to concluded the Africa has the highest cases of child mortality, with all the countries with high cases are in Africa. 