---
title: "Project2_3_Bikram"
author: "Bikram Barua"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project 2_3 (Dataset of stock market in India)
Original Data Reference: <https://www.kaggle.com/souravroy1/stock-market-data/download>


##### Let's load the required libraries in R for data analysis
```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)
library(downloader)
library(stringr)
library(htmlTable)
library(tidyverse)
library(reshape2)

```

<br>

##### Download and read the CSV file to a dataframe object  
##### View the data  

```{r message=FALSE, echo=TRUE, warning=FALSE}

# Import the data from the csv file

stock_market_data <- read.csv("https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/Stock_Market_data.csv", stringsAsFactors = FALSE)

head(stock_market_data, 5) 

```

#### Filter the rows with non-standard Sector values
#### Clean the data, remove # values 

```{r}


stock_market_df <- stock_market_data%>% filter(str_detect(Sector, "^[A-Z]"))  %>%  select(Share,Sector ,Last.Traded.Price, 
 Percentage.Change,High.Price, Low.Price,Yearly.Gainner,PeRatio,New.Pe, W52_High, MarketCap, DividendYield ) %>% arrange(Sector)

stock_market_df <- sapply(stock_market_df,function(x) {x <- gsub("#N/A" ,"", gsub("#DIV/0!","",x)) })

head(stock_market_df, 5)

```

#### Statistical value by Sector
#### Clean the data
#### Sorted by number of companies per Sector with their mean Last traded price 
```{r warning=FALSE, message=FALSE}
stock_market_df <-  as.data.frame(stock_market_df)

stock_market_df$Last.Traded.Price <- as.numeric(stock_market_df$Last.Traded.Price)

summarize_Sector <- stock_market_df %>% group_by( Sector) %>% summarise(Mean_Last_Traded_Price = round(mean(Last.Traded.Price, na.rm=TRUE), 2), Num_Companies = n()) %>% filter(!is.na(mean)) %>% arrange(desc(Num_Companies))

head(summarize_Sector)
```
#### Sort the companies by their Market Cap, display their Yearly Gain %, PE Ratio and Divident Yield
#### Display top 20 companies

```{r warning=FALSE, message=FALSE}

stock_market_df$MarketCap <- as.numeric(stock_market_df$MarketCap)

summarize_MarketCap <- head( stock_market_df %>% filter(!is.na(MarketCap)) %>% arrange(desc(MarketCap)), 20) %>% select(Share, Yearly.Gainner,PeRatio,MarketCap, DividendYield )

htmlTable(summarize_MarketCap)


```

