---
title: "DATA_607_Project2_1"
author: "Bikram Barua"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 607 Project2_1 Homework
##### Global annual GDP growth numbers by country from 1960 - 2020
Data Reference Link: [data.worldbank.org](https://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG)


##### Let's load the required libraries in R for data analysis
```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)
library(downloader)
library(stringr)
library(htmlTable)
library(tidyverse)

```

<br>

##### Download and read the CSV file to a dataframe object  
##### View the data  

```{r message=FALSE, echo=TRUE, warning=FALSE}

# Import the data from the csv file
gdp_data <- read.csv("https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/GDP_by_country.csv", stringsAsFactors = FALSE)

head(gdp_data, 5)  

```

##### Remove unwanted columns

```{r}
gdp_data2  <- gdp_data %>% mutate_if(is.numeric, round, digits=2)
subset_gdp_data <- gdp_data2 %>% select(-'Indicator.Name',-'Indicator.Code')

```

##### Remove the letter 'X' from the column names
##### Rename the first column of the data frame

```{r}
names(subset_gdp_data) <- gsub("X", "", names(subset_gdp_data))

subset_gdp_data <-  subset_gdp_data %>% {colnames(.)[1] = "country"; .}

```


##### Load latitude/longitude data

```{r}

# Import the country latitude/longitude data from the csv file
country_latlng_data <- read.csv("https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/country_latitude_longitude.csv")
head(country_latlng_data)

```

##### Select relevant columns from the country geo dataset
##### Re-order the columns in the dataframe

```{r}

subset_latlng_data <- subset(country_latlng_data, select= c('latitude', 'longitude', 'country'))

col_order <- c("country","latitude","longitude")
new_latlng_data <- subset_latlng_data[, col_order]

```

##### Join the latitude/longitude with GDP columns in the dataframe

```{r}
joined_df <- left_join( new_latlng_data,subset_gdp_data, by ="country")


```
#### Selecting few columns to work with
#### Add the GDP for last 5 years, create a new column with the total GDP
#### Filter out the countries which have not reported their GDP every year

```{r}
### Selecting few columns

sel_df <-  select(joined_df, country,'2016', '2017', '2018', '2019' ,'2020', latitude, longitude) %>%
  mutate(total =  joined_df$'2016' + joined_df$'2017'+  joined_df$'2018'+ joined_df$'2019' + joined_df$'2020') %>% filter(!is.na(total))


```

#### Sort the countries by their total GDP
#### Display the top 10 performing countries from (2016 - 2020)
```{r}
top_10_countries <-  sel_df %>% arrange(desc(total)) %>% head(10) 
knitr::kable(top_10_countries, "html")
```

#### Plot bar graph for the top countries

```{r}
# Basic histogram

ggplot(top_10_countries, aes(x=country, y=total)) + geom_bar(stat="identity") +
  scale_x_discrete(guide = guide_axis(angle = 90)) + NULL
```

#### Mark the top countries on the world map

```{r warning=FALSE, message=FALSE}

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

  
  ggplot(data = world) +
    geom_sf() +
  geom_point(data = top_10_countries, aes(x = longitude, y = latitude), color = "#e60000") +
  geom_text(data= top_10_countries,aes(x=longitude, y=latitude, label=country),
    color = "darkblue", fontface = "bold", size = 3, check_overlap = TRUE)  +
  scale_fill_manual(values = c("#CCCCCC","#e60000")) +
  labs(title = 'Countries with highest GDP') +
  theme(text = element_text(family = "Arial", color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#FFFFF4")
        ,plot.background = element_rect(fill = "#FFFFF4")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 30)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ,legend.position = "none"
        )
```

