---
title: "Tidyverse_Create_Bikram"
author: "Bikram Barua"
date: "10/24/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(gridExtra)
```

# Heart Failure Prediction Dataset

## 11 clinical features for predicting heart disease events.

Source: https://www.kaggle.com/fedesoriano/heart-failure-prediction

<br>

```{r, warning=FALSE, message=FALSE}
heart_failure_predictions <- read_csv("https://raw.githubusercontent.com/baruab/msdsrepo/main/DATA-607/heart_failure_prediction.csv")
#,                  col_names = FALSE)

head(heart_failure_predictions)
```

<br>

##### glimpse () function provides a snapshot of the file in R. It tells number of Rows and Column with listing of the data/types of the columns.

```{r}
glimpse(heart_failure_predictions)
```

<br>

### Give descriptive names to the columns

```{r}
names(heart_failure_predictions) <- c('Age', 'Sex', 'Chest_Pain_Type', 'Resting_Blood_Pressure', 'Cholesterol', 'Fasting_Blood_Sugar', 'Resting_ECG', 'Max_Heart_Rate', 'Excercise_Angina', 'Old_Peak', 'ST_Slope', 'Heart_Disease')



```


<br>


##### List the unique types of chest pains
```{r}
unique(heart_failure_predictions$Chest_Pain_Type)
```

<br>

### Converting the chest_pain acronyms to descriptive names
```{r}

pain_types <- c("ATA" = "Atypical Angina",
                "NAP"= "Non-Anginal Pain",
                "ASY"= "Asymptomatic",
                "TA" = "Typical Angina")
heart_failure_predictions$Chest_Pain_Type <- pain_types[heart_failure_predictions$Chest_Pain_Type]
head(heart_failure_predictions)
```

<br>


##### Barplots for Age, Resting BP and Max Heart Rate

```{r}
g1 <- ggplot(heart_failure_predictions, aes(Age)) + geom_bar() 
g2 <- ggplot(heart_failure_predictions, aes(Resting_Blood_Pressure)) + geom_bar()
g3 <- ggplot(heart_failure_predictions, aes(Max_Heart_Rate)) + geom_bar() 

grid.arrange(g1, g2,g3, nrow=3)  
```

<br>


##### Resting ST-T wave abnormalities have been associated with an increased risk of adverse cardiac events.

```{r}
# Filter by ECG ST abnormality
ST_cases <-filter(heart_failure_predictions,Resting_ECG=="ST",Cholesterol > 200)
```


##### For cases with Resting ST-T wave abnormalities

```{r}

# Heart disease by Age,Resting BP with ST abnormalities
ST_cases %>% ggplot()+
  geom_bar(aes(Age,Resting_Blood_Pressure,fill=Heart_Disease),position="dodge",stat = "identity")+
  theme(axis.text.x=element_text(angle=0))+
  labs(x="Age",y="Resting BP", title="Statistics on heart disease by Age",
       subtitle= "Heart disease by Age,Chest Pain Type, ST condition")


```


<br>


##### For All cases (no filter)


```{r}

# Heart disease by Age,Chest Pain Type, ST condition
heart_failure_predictions %>% ggplot()+
  geom_bar(aes(Age,Resting_Blood_Pressure,fill=Heart_Disease),position="dodge",stat = "identity")+
  theme(axis.text.x=element_text(angle=0))+
  labs(x="Age",y="Resting Blood Pressure", title="Statistics on heart disease by Age",
       subtitle= "Heart disease by Age,Chest Pain Type, ST condition")

```



<br>

##### Fo All cases by Age, Resting BP and Chest Pain Type


```{r}

# Scatter Plot by Age, Resting BP and Chest Pain Type

s1 <- ggplot(heart_failure_predictions, aes( Resting_Blood_Pressure, Age, color = Excercise_Angina)) + 
  geom_point()

s2 <- ggplot(heart_failure_predictions, aes( Max_Heart_Rate, Age, color = Excercise_Angina)) + 
  geom_point()


s3 <- ggplot(heart_failure_predictions, aes( Resting_Blood_Pressure, Age, color = Sex)) + 
  geom_point()

s4 <- ggplot(heart_failure_predictions, aes( Max_Heart_Rate, Age, color = Sex)) + 
  geom_point()


grid.arrange(s1, s2, s3, s4,  nrow = 4)

```







### Conclusion

##### This is to summarise few capabilities of using R Tidyverse to manipulate and interpret using the types of visualisations. Tidyverse can accomodate more continuous and categorical variables to see if there is a relationship or causation exist in a particular situation.