---
title: "DATA_606_Lab1"
author: "Bikram Barua"
date: "8/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 606 Lab
#### Load the required libraries and the data
```{r  load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)

data('arbuthnot', package='openintro')
```

#### Lets take a look at the data
```{r}
arbuthnot
```


#### Glimpse of the same data 
```{r}
glimpse(arbuthnot)
```

#### Let's explore the data
#### Boys baptization information
```{r}
arbuthnot$boys
```

<br>

#### **Exercise 1:**  Extract just the counts of girls baptized
```{r}
arbuthnot$girls
```

#### Total num of girls baptized over the years
```{r}
sum_girls <-  sum(arbuthnot$girls)
sum_girls
```

#### Plot Girls baptization 
```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point()
```

#### Plot the same as Line graph
```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()
```

<br>

#### **Exercise 2:** How would you describe the apparent trend in the number of girls baptized over the years? 
#### Draw a bar plot to represent the trend of girls baptized over the years 
```{r}
# Create Sub-set data
girls_subdata <- subset(arbuthnot, select = c("year","girls"))
p<-ggplot(data=girls_subdata, aes(x=year, y=girls)) +
  geom_bar(stat="identity")
p
```

#### From the above plot, it is clear that as year progressed, we have noticed increase in the count of girls being baptized.

<br>

#### Adding few new columns(total, boys to girls ratio, boys ratio) in the dataframe

```{r}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)
```

<br>

#### **Exercise 3:** Generate a plot of the proportion of boys born over time 
#### Calculate boy proportion as percentage
```{r}
arbuthnot <- arbuthnot %>%
  mutate(boy_percent = (boys / total) * 100 )
b<-ggplot(data=arbuthnot, aes(x=year, y=boy_percent)) +
  geom_bar(stat="identity")
b
```

#### Boys percentage has always been higher than 50, which means more boys were baptized compared to girls over the years


```{r}
arbuthnot %>%
  summarize(min = min(boys), max = max(boys))
```

<br>

####  **Exercise 4:** What years are included in this data set? What are the dimensions of the data frame? What are the variable (column) names?
#### All the years from the original arbuthnot dataset in included. The result data frame is of 1 x 2 dimension. The variable column names are min and max.

<br>

#### **Exercise 5:** How do these counts compare to Arbuthnot’s? Are they of a similar magnitude?
#### These counts are same as Arbuthnot, with similar magnitude.

<br>

#### **Exercise 6:** Make a plot that displays the proportion of boys born over time. What do you see?
```{r}
b<-ggplot(data=arbuthnot, aes(x=year, y=boy_ratio)) +
  geom_line()
b
```

#### Based on the above graph of boys propotion compared to girls, it is clear that during the time period the boys population was higher compared to the girls.

<br>

#### **Exercise 7:** In what year did we see the most total number of births in the U.S.?
```{r}
sorted_total <- arrange(arbuthnot, desc(total))
head(sorted_total, 1)
```

#### The year with the most total number of births in US was 1705

