---
title: "DATA607_Week2_Assignment"
author: "Bikram Barua"
date: "9/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### DATA 607 Assignment - SQL and R

##### Choose six popular movies, which your freinds, co-workers might have watched.
##### Gather the ratings from your friends and try to analyse the result

##### Used MySQL database as a system of record for the above purpose.
##### Created few tables: movies, users and user_ratings to hold the relevant data

##### Let's load the required libraries in R for data analysis
```{r load-packages, message=FALSE, warning=FALSE}
library(dplyr)
library(sqldf)
library(RODBC)
library(DBI)
library(odbc)

library(downloader)
library(htmlTable)
```


<br>

#### Connect to the MySQL database
#### List the movies
```{r }
mydb = dbConnect(odbc(), "MySQL")
                
movies_dt <- dbGetQuery(mydb, "select movie_name from movies")
htmlTable(movies_dt)
```

<br>

#### Join multiple tables to get data in one frame
#### Filter the rows with missing rating data
```{r}
ratings_dt <- dbGetQuery(mydb, "select movie_name, username, userrole, rating
               from user_ratings a
                inner join users b on a.user_id = b.userid 
                inner join movies c on a.movie_id = c.id
               where rating > -1", row.names=FALSE)
htmlTable(head(ratings_dt, 20))
```

<br>


#### Group and sum the ratings by movie names 

```{r}
library(sqldf)

ratings_groupByMovies <- sqldf( "SELECT movie_name, sum(rating)  
                                 FROM ratings_dt
                                   group by movie_name
                                order by sum(rating) desc", row.names=TRUE)

head(ratings_groupByMovies)
```

#### Highest rated movie by friends and co-workers
```{r}
ratings_groupByRole <- sqldf("SELECT movie_name, userrole, sum(rating)  
                                 FROM ratings_dt
                                   group by movie_name, userrole
                                order by movie_name", row.names=TRUE)

htmlTable(head(ratings_groupByRole, 2))
```

<br>

#### Count the users who rated the movies in the survey based on their role
```{r}
countByUserRole<- sqldf("SELECT movie_name,  
                  SUM(
                      CASE WHEN userrole = 'Friend' THEN 1 ELSE 0 END
                    ) As 'Num_Friends_Rated',
                  SUM(
                      CASE WHEN  userrole = 'Co-Worker' THEN 1 ELSE 0 END
                    ) As 'Num_Co-Workers_Rated'
                   FROM ratings_dt
                group by movie_name")

htmlTable(countByUserRole)
```
<br>

#### The users were later asked a survey on various categories about the movies
#### Get survey data from csv file
```{r echo=FALSE, warning=FALSE, message=FALSE}

url <- "https://raw.githubusercontent.com/baruab/msdsrepo/main/Movie_user_ratings.csv"

download_file <- "Movie_user_ratings.csv" # name of the file on the local machine after download

# the file will be downloaded to the working directory
downloader::download(url, download_file)

# read the population dataset into a dataframe

movie_user_survey <- read.csv(download_file, header=TRUE, stringsAsFactors = FALSE)
head(movie_user_survey)
```


```{r}

combined_df <- 
        inner_join(ratings_dt, movie_user_survey, by=c("movie_name", "username"))             
htmlTable(head(arrange(combined_df, movie_name, username), 15))

```




